Discrete Quantum Theories and Computing

### 2019-04-17 dissertation.tex
dissertation.lyx "Chapter 6 Conclusion and Further Discussion" Updating...

Conventional quantum theory is defined over the continuum of complex
numbers, but each complex number can only be identified with unbounded
resources. To explore alternative versions of quantum theory incorporating
our limitation of distinguishability, two types of discrete quantum
theories have been described in this thesis: quantum theories and
computing over finite fields and quantum interval-valued probability
measures. Examining the computational and physical consequences of
such frameworks can yield new insights into not only the subtle properties
of conventional quantum theory but also the power and capacity of
quantum computing.

The theories over finite fields starting with unrestricted discrete
fields (modal theories), then advancing to a more reasonable framework
based on complexifiable discrete fields (discrete quantum theory I),
which lacks a notion of probability and supports unnaturally efficient
deterministic quantum algorithms. A still more plausible discrete
theory (discrete quantum theory II) is proposed, where conventional
quantum computing and conventional quantum theory emerge in a local
sense.

\todo{Clean up until here\ldots}

....

Before discussing the further extension, let's summarize what we have
done previously. In Chapter~\ref{chap:Conventional-Quantum-Theory},
we reviewed what was established in CQT. In particular, Sec.~\ref{CQC1qubitBloch.sec}
studied the geometrical structure of quantum states to prepare replacing
the conventional Hilbert space~$\mathComplex^{D}$ or projective
Hilbert space~$\CP{D-1}$ by its discrete analog in Chapter~\ref{chap:Unrestricted-Finite-Fields}
and~\ref{chap:DQT=000026DQC}. In Chapter~\ref{chap:Unrestricted-Finite-Fields},
the field of complex numbers~$\mathComplex$ is replaced by unrestricted
finite fields~$\Fq$. Although this modal theory is

, and a valid pure state can be represented as any non-zero vector
in $\Fq^{D}$ and a measure scheme could be defined which only predicts
whether an measurement result is possible or not. \todo{Keep writing\ldots}

@@

\begin{comment}
When people proved the original Gleason theorem, people usually exploited
the geometrical structure of real 3-dimensional Hilbert space \cite{gleason1957,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
Since our finite-precision extension of the Gleason theorem only applies
on a class of QIVPMs, we might want to ask how to modify these geometrical
arguments to have a Gleason-type theorem for general QIVPMs. We will
further study the tensor product structure among QIVPMs which is essential
for defining product and entangled states and serves the basis to
discuss quantum nonlocality~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and quantum computing with QIVPMs. Finally, we want to improve the
discrete quantum theories to consider QIVPMs over finite fields in
future research.
\end{comment}
As we discussed in the previous section, Thm.~\ref{thm:Finite-precision-Gleason}
only applies to the QIVPMs constructed by composing an interval map
with a conventional quantum probability measure, and the states consistent
with the composite QIVPM collapse to a single state as the maximum
length of intervals in $\mathscr{I}$, $\left\Vert \mathscr{I}\right\Vert $,
shrinks to $0$. In contrast, the globally inconsistent QIVPM defined
in Table~\ref{tab:non-Born-QIVPM} has the least sharp range~$\mathscr{I}_{0}$
with $\left\Vert \mathscr{I}_{0}\right\Vert =1$. This suggests a
possibility that shrinking the length $\left\Vert \mathscr{I}\right\Vert $
might help to regularize general QIVPMs, and it is natural to ask
whether there is a short enough length~$\varepsilon$ such that QIVPMs
mapping to intervals not longer than $\varepsilon$ always have non-empty
cores.

>>>>

Conventional quantum theory is based on the continuum of complex numbers,
but we cannot distinguish two arbitrary complex numbers without unbounded
resources. To explore alternative versions of quantum theory incorporating
our limitation of distinguishability, two types of discrete quantum
theories were described: \emph{quantum theories and computing over
finite fields} and \emph{quantum interval-valued probability measures
(QIVPMs)}. Examining the computational and physical consequences of
such frameworks could yield new insights into not only the subtle
properties of conventional quantum theory but also the power and capacity
of quantum computing.

The theories over finite fields started with unrestricted discrete
fields (Chapter~\ref{chap:Unrestricted-Finite-Fields}), then advanced
to a more reasonable framework based on complexifiable discrete fields
(Sec.~\ref{discretequantumtheoryI} to~\ref{discretequantumcomputingI}),
which lacked a notion of probability and supported unnaturally efficient
deterministic quantum algorithms. A still more plausible discrete
theory with cardinal probabilities was proposed (Sec.~\ref{sec:Discrete-Quantum-Theory-(II)}
and~\ref{discretequantumcomputingII}), where conventional quantum
theory and computing emerged in a local sense, but lacking arithmetic
operations among cardinal probabilities still posed difficulty to
define expectation values. Since the axiomatic approach looked unlikely
to provide ``sensible'' real-valued probability measures over finite
fields (Sec.~\ref{sec:Toward-IVPM}), we shifted our attention to
directly embed our limitation of distinguishability into the theory
to define QIVPM.

\begin{figure}[b]
\centering{}$\xymatrix{\textrm{Classical Probability Measure}\ar[d]_{\textrm{blur probability}}\ar[r]^{\textrm{glue events}} & \textrm{Quantum Probability Measure}\ar[d]^{\textrm{blur probability}}\\
\textrm{Classical IVPM}\ar[r]_{\textrm{glue events}} & \textrm{QIVPM}
}
$\caption{\label{fig:commutative-diagrams}QIVPMs inherited from both quantum
probability measures and classical IVPMs.\todo{The first arrow is
too short!}}
\end{figure}
As a natural extension of both convention quantum probability measures
and classical interval-valued probability measures (IVPMs) illustrated
in Fig.~\ref{fig:commutative-diagrams}, QIVPMs inherited definitions
and properties from the both sides. While the expectation values of
QIVPMs could be pulled back to the classical cases and consistent
with those of quantum probability measures in the infinitely precise
limit (Sec.~\ref{sec:Interval-Uncertainty}), foundational concepts
in quantum mechanics, such as the Kochen-Specker and Gleason theorems,
extended to QIVPMs in subtle ways. By carefully specifying experimental
uncertainties, we established rigorous bounds on the validity of the
Kochen-Specker theorem (Sec.~\ref{sec:Kochen-Specker}). While there
is a QIVPM not consistent with Gleason's unique state~$\rho$ on
all projectors, we constructed a class of QIVPMs for which the original
Gleason theorem could be recovered asymptotically (Sec.~\ref{sec:Gleason}).

In the following further discussion, we will briefly explain why we
can only recover Gleason's on a class of QIVPM, the possibility to
further build a computational model over QIVPMs, and the possibility
to combine two approaches to consider quantum interval-valued probability
measures over finite fields.

@@

\begin{comment}
When people proved the original Gleason theorem, people usually exploited
the geometrical structure of real 3-dimensional Hilbert space \cite{gleason1957,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
Since our finite-precision extension of the Gleason theorem only applies
on a class of QIVPMs, we might want to ask how to modify these geometrical
arguments to have a Gleason-type theorem for general QIVPMs. We will
further study the tensor product structure among QIVPMs which is essential
for defining product and entangled states and serves the basis to
discuss quantum nonlocality~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and quantum computing with QIVPMs. Finally, we want to improve the
discrete quantum theories to consider QIVPMs over finite fields in
future research.
\end{comment}
As we discussed in Sec.~\ref{sec:Gleason}, Thm.~\ref{thm:Finite-precision-Gleason}
only applies to the QIVPMs constructed by composing an interval map
with a conventional quantum probability measure, and the states consistent
with the composite QIVPM collapse to a single state as the maximum
length of intervals in $\mathscr{I}$, $\left\Vert \mathscr{I}\right\Vert $,
shrinks to $0$. In contrast, the globally inconsistent QIVPM defined
in Table~\ref{tab:non-Born-QIVPM} has the least sharp range~$\mathscr{I}_{0}$
with $\left\Vert \mathscr{I}_{0}\right\Vert =1$. This suggests a
possibility that shrinking the length $\left\Vert \mathscr{I}\right\Vert $
might help to regularize general QIVPMs, and it is natural to ask
whether there is a short enough length~$\varepsilon$ such that QIVPMs
mapping to intervals not longer than $\varepsilon$ always have non-empty
cores.
